---
title: "Gender and College Majors"
author: "Kellie Haddon"
institute: "SIS 750: Data Analysis"
output: 
  beamer_presentation:
   theme: "default"
   colortheme: "dove"
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center')

## Packages
library(tidyverse)
library(haven)
library(forcats)
library(scales)
library(extrafont)
library(showtext)
library(knitr)

## Load data
df = read_csv('recent-grads.csv')
```

## **Data Description**
Data Sources: American Community Survey 2010-2012 Public Use Microdata Series & FiveThirtyEight

### **Sample**
* Individual-level survey data
* Full-time, year-round workers
* Undergraduate degree majors
* Ages: <28  

```{r, descriptives}
Variable = c('Total', 'Men', 'Women', 'Sample', 'Majors', 'Major Categories')
n = c('6,771,654', '2,876,426', '3,895,228', '61,602', '173', '16')

df2 = tibble(Variable, n)

kable(df2)
```

## **Share of Women by Major Category**
Percentage of full time workers who are women in each major category
```{r, fig.dim=c(4.5,3), analysis1}
df3 <- df %>%
  group_by(Major_category) %>%
  summarise(ShareWomen = mean(ShareWomen, na.rm = TRUE))

df3 %>%
  ggplot(aes(x = reorder(Major_category, +ShareWomen), y = ShareWomen)) + 
  geom_col(fill = '#41afaa') +
  coord_flip() +
  labs(x = ' ', y = 'Share of Women', caption = ' ') +
  geom_text(aes(label = paste0(round(ShareWomen * 100, 1), "%")), 
            position = position_stack(vjust = 0.5), color = "white", size = 2.5, face = 'bold') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(color = 'black', size = 8),
        axis.text.y = element_text(color = 'black', size = 8, margin = margin(t = 0, r = -5, b = 0, l = 0)),
        axis.title.x = element_text(color = 'black', size = 9),
        axis.title.y = element_text(color = 'black', size = 9))
```

## **Share of Women vs. Median Earnings**
Majors plotted by the share of full time workers who are women and the median earnings of workers in each major
```{r, fig.dim=c(4.5,2.75), analysis2}
df = df %>%
  mutate(STEM = (case_when(Major_category == 'Agriculture & Natural Resources' ~ 'STEM',
                           Major_category == 'Arts' ~ 'Non-STEM',
                           Major_category == 'Biology & Life Science' ~ 'STEM',
                           Major_category == 'Business' ~ 'Non-STEM',
                           Major_category == 'Communications & Journalism' ~ 'Non-STEM',
                           Major_category == 'Computers & Mathematics' ~ 'STEM',
                           Major_category == 'Education' ~ 'Non-STEM',
                           Major_category == 'Engineering' ~ 'STEM',
                           Major_category == 'Health' ~ 'Non-STEM',
                           Major_category == 'Humanities & Liberal Arts' ~ 'Non-STEM',
                           Major_category == 'Industrial Arts & Consumer Services' ~ 'Non-STEM',
                           Major_category == 'Interdisciplinary' ~ 'Non-STEM',
                           Major_category == 'Law & Public Policy' ~ 'Non-STEM',
                           Major_category == 'Physical Sciences' ~ 'STEM',
                           Major_category == 'Psychology & Social Work' ~ 'Non-STEM',
                           Major_category == 'Social Science' ~ 'Non-STEM')))

df$STEM = as_factor(df$STEM)

df %>%
  ggplot(aes(x = ShareWomen, y = Median)) +
  geom_point(color = '#af4b91', alpha = 0.2, size = 3) +
  geom_smooth(se = F, color = '#af4b91') +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(x = 'Share of Women', 
       y = 'Median Earnings') +
  theme_minimal() +
  theme(axis.text.x = element_text(color = 'black',size = 8),
        axis.text.y = element_text(color = 'black',size = 8),
        axis.title.x = element_text(color = 'black',size = 9),
        axis.title.y = element_text(color = 'black',size = 9))
```

## **Share of Women vs. Median Earnings by Degree Type**
Majors plotted by the share of full time workers who are women and the median earnings of workers in each major separated by degree type (STEM vs. non-STEM)
```{r, fig.dim=c(4.25,2.25), analysis3}
    # sub-group STEM
df %>%
  ggplot(aes(x = ShareWomen, y = Median, color = STEM)) +
  geom_point(alpha = 0.2, size = 3) +
  geom_smooth(se = F) +
  facet_wrap(~ STEM) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::dollar_format()) +
  scale_color_manual(values = c('#d7642c', '#e6a532')) +
  labs(x = 'Share of Women', 
       y = 'Median Earnings') +
  guides(color = FALSE) +
  theme_minimal() +
  theme(strip.text.x = element_text(color = 'black',size = 8, face = 'bold'),
        axis.text.x = element_text(color = 'black',size = 8),
        axis.text.y = element_text(color = 'black',size = 8),
        axis.title.x = element_text(color = 'black',size = 9),
        axis.title.y = element_text(color = 'black',size = 9),
        panel.margin = unit(1, "lines"))
```
