---
title: "Top Earning College Majors"
author: "Kellie Haddon"
output: 
  beamer_presentation:
   theme: "Copenhagen"
   colortheme: "seagull"
   fig_width: 4
   fig_height: 3
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center')

## Packages
library(tidyverse)
library(haven)
library(forcats)
library(scales)
library(extrafont)
library(showtext)
library(knitr)

## Load data
df = read_csv('recent-grads.csv')
```

## Data Description
Data Sources: American Community Survey 2010-2012 Public Use Microdata Series & FiveThirtyEight

### Sample  
* Ages: <28  
* Full-time, year-round

```{r, descriptives}
Variable = c('Total', 'Men', 'Women', 'Sample', 'Majors', 'Major Categories')
n = c('6,771,654', '2,876,426', '3,895,228', '61,602', '173', '16')

df2 = tibble(Variable, n)

kable(df2)
```

## Top 100 Earning Major Categories

```{r, analysis1}
df %>%
  filter(Rank < 101) %>%
  ggplot(aes(fct_infreq(Major_category))) +
  geom_bar(fill = '#41afaa', alpha = 0.6, color = '#41afaa') +
  labs(x = ' ', y = ' ', caption = ' ', title = '') +
  geom_text(stat='count', aes(label = ..count..), vjust = -0.7, size = 2, color = 'black') +
  scale_y_continuous(limits = c(0, 35)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, vjust = 1.15, hjust = 1, color = 'black', size = 6),
        panel.grid = element_blank(),
        axis.text.y = element_blank())
```

## Share of Women vs. Median Earnings

```{r, analysis2, width = 8, height = 5}
df = df %>%
  mutate(STEM = (case_when(Major_category == 'Agriculture & Natural Resources' ~ 'STEM',
                           Major_category == 'Arts' ~ 'Non-STEM',
                           Major_category == 'Biology & Life Science' ~ 'STEM',
                           Major_category == 'Business' ~ 'Non-STEM',
                           Major_category == 'Communications & Journalism' ~ 'Non-STEM',
                           Major_category == 'Computers & Mathematics' ~ 'STEM',
                           Major_category == 'Education' ~ 'Non-STEM',
                           Major_category == 'Engineering' ~ 'STEM',
                           Major_category == 'Health' ~ 'Non-STEM',
                           Major_category == 'Humanities & Liberal Arts' ~ 'Non-STEM',
                           Major_category == 'Industrial Arts & Consumer Services' ~ 'Non-STEM',
                           Major_category == 'Interdisciplinary' ~ 'Non-STEM',
                           Major_category == 'Law & Public Policy' ~ 'Non-STEM',
                           Major_category == 'Physical Sciences' ~ 'STEM',
                           Major_category == 'Psychology & Social Work' ~ 'Non-STEM',
                           Major_category == 'Social Science' ~ 'Non-STEM')))

df$STEM = as_factor(df$STEM)

df %>%
  ggplot(aes(x = ShareWomen, y = Median)) +
  geom_point(color = '#41afaa', alpha = 0.6) +
  geom_smooth(se = F, color = '#41afaa') +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(x = 'Share of Women', 
       y = 'Median Earnings') +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))
```

## Share of Women vs. Median Earnings by Degree Type

```{r, analysis3}
    # sub-group STEM
df %>%
  ggplot(aes(x = ShareWomen, y = Median, color = STEM)) +
  geom_point(alpha = 0.6) +
  geom_smooth(se = F) +
  facet_wrap(~ STEM) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::dollar_format()) +
  scale_color_manual(values = c('#d7642c', '#af4b91')) +
  labs(x = 'Share of Women', 
       y = 'Median Earnings') +
  guides(color = FALSE) +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 8, face = 'bold'),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))
```
